Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-07 ms.tex

\item Stellar wind heating, supernovae, and AGN feedback depend explicitely on the SMBH mass as well as the SFH in the galactic nuclear region.  A young starburst of age $\lesssim 10$ Myr produces mass and energy injection dominated by young stellar winds ($v_w \sim 1000$ km s$^{-1}$), while for exclusively old stellar populations  mass input is dominated by AGB winds and heating is dominated by AGN feedback and supernovae.  Ongoing, continuous star formation presents a hybrid situation, with energy input usually dominated by young stars and mass input dominated by the old stars (Appendix \ref{app:windheat}).  Galactic nuclei with $\Mbh \lesssim 10^8\Msun$ heated according to their average SFHs (as derived from cosmological simulations) receive a stellar wind heating of $v_w\sim 800$ km s$^{-1}$, although this number can be somewhat lower if the star formation is intermittent (Fig.~\ref{fig:NickPlot2}).

\item Type Ia SNe only provide a continuous heating source exterior to the Ia radius $\rIa$ (eq.~[\ref{eq:rIa}]) where the time between subsequent SNe exceeds the dynamical timescale.  Average SFH heating results in $\rIa < \rs$, except for the most massive galaxies.  Nevertheless, even if other sources of heating are low enough that $\rIa > \rs$, the stagnation radius is unlikely to exceed $\rIa$ due to Ia heating.  SN Ia thus cap the SMBH accretion rate at the value $\dot{M}_{Ia}$ (eq.~[\ref{eq:MdotIa}]), except possibly in high mass elliptical galaxies with large break radii, for which $\zeta(v_w^{\rm Ia}) \gtrsim \zeta_c$.

\item Stellar wind heating due to the average SFH is sufficient to result in thermally-stable, steady-state accretion.  The resulting average Eddington-normalized accretion rate is predicted to increase weakly with $M_{\bullet}$ (Fig.~\ref{fig:miller}), in tension with the (also weak) downsizing trend of measured nuclear X-ray luminosities of early-type galaxies (e.g., \citealt{Miller+15}).  However, given the large uncertainties in translating predictions for the gas inflow rate on parsec scales to the disk X-ray luminosity, these differences may not be significant.

Thermally stable accretion models can reproduce scatter in nuclear X-ray luminosities at fixed $\Mbh$ of the observed 2-4 orders of magnitude from a combination of (1) intrinsic scatter in the SMBH scaling relations; (2) stochastic variation in SFH from galaxy to galaxy, or from time to time due to burstiness in the SFH (e.g.~\ref{fig:NickPlot2}); (3) Variations in the amount of inflowing mass which is lost to small scale outflows from the SMBH accretion disk, due to variations in the angular momentum of the accreting gas.

\item Only for relatively high X-ray radiative efficiencies similar to the maximum values predicted by ADAF models are the inferred accretion rates of quiescent early-type galaxies below the maximum value allowed for thermally-stable accretion (Fig.~\ref{fig:miller}).  Massive galaxies are also particularly susceptible to thermal instability, given the large stagnation radii that can result from the deep potentials and large break radii of these (mainly core) galaxies.  It is thus possible that the current X-ray detected sample is instead comprised mainly of nuclei undergoing non-steady accretion as a part of local thermal instabilities.  In this case a separate branch of low $L_x$ (undetected) thermally-stable nuclei may exist, especially at low $M_{\bullet}$.  Such a dichotomy challenges the practice of extrapolating the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$ to constrain the SMBH occupation fraction in low mass galaxies (\citealt{Miller+15}).

  \item Low mass black holes grow in the low redshift universe over timescales comparable to the Hubble time (\citet{Heckman+04}).  This accretion rates so required are too high to be consistent with the maximum allowed for steady-state, thermally stable accretion (Fig.~\ref{fig:bh_growth}).  Low-$M_{\bullet}$ growth and AGN activity must instead be driven by gas supplied externally to the nuclear region, e.g. due to galaxy mergers or thermal instabilities on larger, galactic scales (e.g.~\citealt{Voit+15}).

>>>>

\item Stellar wind heating, supernovae, and AGN feedback depend explicitely on the SMBH mass as well as the SFH in the galactic nuclear region.  A young starburst of age $\lesssim 10$ Myr produces mass and energy injection dominated by young stellar winds ($v_w \sim 1000$ km s$^{-1}$), while for exclusively old stellar populations  mass input is dominated by AGB winds and heating is dominated by AGN feedback and supernovae.  Ongoing, continuous star formation presents a hybrid situation, with energy input usually dominated by young stars and mass input dominated by the old stars (Appendix \ref{app:windheat}).  Galactic nuclei with $\Mbh \lesssim 10^8\Msun$ heated according to their average SFHs (as derived from cosmological simulations) receive a stellar wind heating of $v_w \sim 700$ km s$^{-1}$, although this number can be somewhat lower if the star formation is intermittent (Fig.~\ref{fig:NickPlot2}).

\item Type Ia SNe only provide a continuous heating source exterior to the Ia radius $\rIa$ (eq.~[\ref{eq:rIa}]) where the time between subsequent SNe exceeds the dynamical timescale.  Heating from the average SFH results in $\rIa < \rs$, except for the most massive galaxies (Fig.~\ref{fig:vwSources}).  Nevertheless, even if other sources of heating are low enough for $\rs$ to approach $\rIa$, the SN Ia heating supplied when $\rs \gtrsim \rIa$ is sufficiently large that $\rs$ is unlikely to greatly exceed $\rIa$.  SNe Ia thus cap the SMBH accretion rate at the value $\dot{M}_{\rm Ia}$ (eq.~[\ref{eq:eddr_Ia}]), except possibly in high mass elliptical galaxies with large break radii, for which $\zeta(v_w^{\rm Ia}) \gtrsim \zeta_c$.

\item Stellar wind heating from the average SFH is sufficient to permit thermally-stable, steady-state accretion.  However, the resulting average Eddington-normalized accretion rate is predicted to increase weakly with $M_{\bullet}$ (Fig.~\ref{fig:miller}), in tension with the (also weak) downsizing trend of measured nuclear X-ray luminosities of early-type galaxies (e.g., \citealt{Miller+15}).  However, these differences may not be discriminating given the huge uncertainty in translating the predicted gas inflow rate on parsec scales to the disk X-ray luminosity.

Thermally stable accretion models can reproduce the observed scatter in nuclear X-ray luminosities at fixed $\Mbh$ of 2-4 orders of magnitude due to combination of (1) intrinsic scatter in the SMBH scaling relations; (2) stochastic variation in SFH from galaxy to galaxy, or from time to time due to burstiness in the SFH (e.g.~\ref{fig:NickPlot2}); (3) Variations in the amount of inflowing mass which is lost to small scale outflows from the SMBH accretion disk, due to variations in the angular momentum of the accreting gas (e.g.~\citealt{Li+13}).

\item Only for high X-ray radiative efficiencies similar to the maximum values predicted by ADAF models are the inferred accretion rates of quiescent early-type galaxies below the maximum value allowed for thermally-stable accretion $\dot{M}_{\rm TI}$ (Fig.~\ref{fig:miller}).  It is thus well possible that the current sample of X-ray detections is instead comprised mainly of nuclei undergoing non-steady accretion due to (local) thermal instabilities.  In this case there may exist a separate branch of low $L_x$ (undetected) thermally-stable nuclei, especially at low $M_{\bullet}$.  Such a bimodal population would call into question extrapolating the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$ to constrain the SMBH occupation fraction in low mass galaxies (\citealt{Miller+15}).

  \item Low mass black holes grow in the low redshift universe over timescales comparable to the Hubble time (\citet{Heckman+04}).  This accretion rates so required are too high to be consistent with the value predicted for the average SFH history and the maximum allowed for steady-state, thermally stable accretion (Fig.~\ref{fig:bh_growth}).  Low-$M_{\bullet}$ growth and AGN activity must instead be driven by a supply of gas external to the nucleaus, e.g. due to galaxy mergers or thermal instabilities on larger, galactic scales (e.g.~\citealt{Voit+15}).
